### 2018-09-25 documentation/authors.tex
Added LaTeX file containing authors

\author{B.~Matusch}
\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{C.~Amole}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{M.~Ardid}
\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}

\author{I.~J.~Arnquist}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{D.~M.~Asner}
%\affil{Brookhaven National Laboratory, Upton, New York 11973, USA}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{D.~Baxter}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
\affil{Enrico Fermi Institute, KICP and Department of Physics, 
University of Chicago, Chicago, Illinois 60637, USA}
%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{E.~Behnke}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{P.~Bhattacharjee}
%\affil{Astroparticle Physics and Cosmology Division, Saha Institute of Nuclear Physics, Kolkata, India}

%\author{H.~Borsodi}
%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{M.~Bou-Cabo}
%\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}

\author{M.~Bressler}
\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}

\author{B.~Broerman}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

%\author{P.~Campion}
%\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}

\author{G.~Cao}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{C.~J.~Chen}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}

\author{U.~Chowdhury}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{K.~Clark}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
%\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
%\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{J.~I.~Collar}
\affil{Enrico Fermi Institute, KICP and Department of Physics,
University of Chicago, Chicago, Illinois 60637, USA}

\author{P.~S.~Cooper}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{M.~Crisler}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{G.~Crowder}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{N.A.~Cruz-Venegas}
\affil{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}

\author{C.~E.~Dahl}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{M.~Das}
\affil{Astroparticle Physics and Cosmology Division, Saha Institute
of Nuclear Physics, Kolkata, India}

\author{S.~Fallows}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{J.~Farine}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{I.~Felis}
\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}

\author{R.~Filgas}
\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}

\author{F.~Girard}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{G.~Giroux}
% \email{ggiroux@owl.phy.queensu.ca}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{J.~Hall}
%\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{C.~Hardy}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{O.~Harris}
\affil{Northeastern Illinois University, Chicago, Illinois 60625, USA}

\author{E.~W.~Hoppe}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{M.~Jin}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}

\author{L.~Klopfenstein}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{C.~B.~Krauss}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{M.~Laurin}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{I.~Lawson}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{A.~Leblanc}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{I.~Levine}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{W.~H.~Lippincott}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{F.~Mamedov}
\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}

\author{D.~Maurya}
\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}

\author{P.~Mitra}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{C.~Moore}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{T.~Nania}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{R.~Neilson}
\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}

\author{A.~J.~Noble}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{P.~Oedekerk}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{S.~Olson}
%\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{A.~Ortega}
\affil{Enrico Fermi Institute, KICP and Department of Physics,
University of Chicago, Chicago, Illinois 60637, USA}

\author{M.-C.~Piro}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{A.~Plante}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{R.~Podviyanuk}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{S.~Priya}
\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}

\author{A.~E.~Robinson}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

%\author{A.~Roeder}
%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{R.~Rucinski}
%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{S.~Sahoo}
\affil{Astroparticle Physics and Cosmology Division, Saha Institute
of Nuclear Physics, Kolkata, India}

\author{O.~Scallon}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{S.~Seth}
\affil{Astroparticle Physics and Cosmology Division, Saha Institute
of Nuclear Physics, Kolkata, India}

\author{A.~Sonnenschein}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{N.~Starinski}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{I.~\v{S}tekl}
\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}

\author{T.~Sullivan}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{F.~Tardif}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{E.~V\'azquez-J\'auregui}
\affil{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{N.~Walkowski}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{J.~Wells}
%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{U.~Wichoski}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{Y.~Yan}
\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}

\author{V.~Zacek}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{J.~Zhang}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}