### 2018-11-23 documentation/authors.tex
Updated authors

\author[1]{B.~Matusch\thanks{Corresponding: \url{brendon.matusch@snolab.ca} or \url{analysis@picoexperiment.com}}}
\affil[1]{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author[2]{C.~Amole}
\affil[2]{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author[3]{M.~Ardid}
\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}

\author[4]{I.~J.~Arnquist}
\affil[4]{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author[4]{D.~M.~Asner\thanks{now at Brookhaven National Laboratory}}

\author[5, 6]{D.~Baxter}
\affil[5]{Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
\affil[6]{Enrico Fermi Institute, KICP and Department of Physics, University of Chicago, Chicago, Illinois 60637, USA}

\author[7]{E.~Behnke}
\affil[7]{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author[8]{M.~Bressler}
\affil[8]{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}

\author[2]{B.~Broerman}

\author[2]{G.~Cao}

\author[5]{C.~J.~Chen}

\author[2]{U.~Chowdhury\thanks{now at Canadian Nuclear Laboratories}}

\author[2]{K.~Clark}

\author[6]{J.~I.~Collar}

\author[9]{P.~S.~Cooper}
\affil[9]{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author[12]{C.~B.~Coutu}

\author[4]{C.~Cowles}

\author[4, 9]{M.~Crisler}

\author[2]{G.~Crowder}

\author[12]{N.~A.~Cruz-Venegas}
\affil[10]{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}

\author[5, 9]{C.~E.~Dahl}

\author[11]{M.~Das}
\affil[11]{Astroparticle Physics and Cosmology Division, Saha Institute of Nuclear Physics, Kolkata, India}

\author[12]{S.~Fallows}
\affil[12]{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author[13]{J.~Farine}
\affil[13]{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author[3]{I.~Felis}

\author[14]{R.~Filgas}
\affil[14]{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}

\author[13, 15]{F.~Girard}
\affil[15]{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author[2]{G.~Giroux}

\author[1]{J.~Hall}

\author[2]{C.~Hardy}

\author[16]{O.~Harris}
\affil[16]{Northeastern Illinois University, Chicago, Illinois 60625, USA}

\author[13]{T.~Hillier}

\author[4]{E.~W.~Hoppe}

\author[4]{C.~M.~Jackson}

\author[5]{M.~Jin}

\author[7]{L.~Klopfenstein}

\author[12]{C.~B.~Krauss}

\author[15]{M.~Laurin}

\author[1, 13]{I.~Lawson}

\author[13]{A.~Leblanc}

\author[7]{I.~Levine}

\author[13]{C.~Licciardi}

\author[9]{W.~H.~Lippincott}

\author[4]{B.~Loer}

\author[14]{F.~Mamedov}

\author[12]{P.~Mitra}

\author[2]{C.~Moore}

\author[7]{T.~Nania}

\author[8]{R.~Neilson}

\author[2]{A.~J.~Noble}

\author[7]{P.~Oedekerk}

\author[6]{A.~Ortega}

\author[12]{M.-C.~Piro}

\author[15]{A.~Plante}

\author[13]{R.~Podviyanuk}

\author[17]{S.~Priya}
\affil[17]{Materials Research Institute, Penn State, University Park, Pennsylvania 16802, USA}

\author[15]{A.~E.~Robinson}

\author[11]{S.~Sahoo}

\author[13]{O.~Scallon}

\author[11]{S.~Seth}

\author[9]{A.~Sonnenschein}

\author[15]{N.~Starinski}

\author[14]{I.~\v{S}tekl}

\author[2]{T.~Sullivan}

\author[15]{F.~Tardif}

\author[10, 13]{E.~V\'azquez-J\'auregui}

\author[7]{N.~Walkowski}

\author[13]{E.~Weima}

\author[13]{U.~Wichoski}

\author[4]{K.~Wierman}

\author[17]{Y.~Yan}

\author[15]{V.~Zacek}

\author[5]{J.~Zhang\thanks{now at Argonne National Laboratory}}